Growing libraries of concepts with wake-sleep program induction

### 2018-07-24 talk.tex
talk improvements

\begin{frame}{EC --- Wake}
  \only<1>{
    \includegraphics[width=11cm]{figures/teachDC/.inkslides-ec/slide-0.pdf}
  }
  \only<2>{
    \includegraphics[width=11cm]{figures/teachDC/.inkslides-ec/slide-1.pdf}
  }
  \only<3>{
    \includegraphics[width=11cm]{figures/teachDC/.inkslides-ec/slide-2.pdf}
  }
  \only<4>{
    \includegraphics[width=11cm]{figures/teachDC/.inkslides-ec/slide-3.pdf}
  }
  \only<5>{
    \includegraphics[width=11cm]{figures/teachDC/.inkslides-ec/slide-4.pdf}
  }

\begin{frame}{DreamCoder --- Sleep-G}
  \only<1>{
    \includegraphics[width=11cm]{figures/teachDC/.inkslides-sleep-g/slide-4.pdf}
  }
\end{frame}
\begin{frame}{DreamCoder --- Sleep-R-1}
  \only<1>{
    \includegraphics[width=11cm]{figures/teachDC/.inkslides-sleep-r-1/slide-4.pdf}
  }
\end{frame}
\begin{frame}{DreamCoder --- Sleep-R-2}
  \only<1>{
    \includegraphics[width=11cm]{figures/teachDC/.inkslides-sleep-r-2/slide-4.pdf}
  }
\end{frame}

@@

\begin{frame}[t]{DSL learning as Bayesian inference}
  %  \includegraphics[width = 11cm]{figures/animation/EC.eps}
\centering  \begin{tikzpicture}[scale=1.3,line width=0.5mm]

@@

\end{frame}
\begin{frame}[t]{Wake: Problem solving}
  \includegraphics[width = 11cm]{figures/animation/Wake.eps}

  \vfill
  \only<1>{
  Recognition model: Neural net.\\
  Red: Bottom-up inference.\\
  Gray: Observed.\\
  White: Latent.\\
  Boxed (plate): Repeated.\\
  }
  \only<2>{
  \begin{tikzpicture}[xscale=1.25,line width=1mm]
    \node at (0,0) (d){DSL};
    \node at ([yshift = -2cm]d) (t){$\text{Task}$};

    \node at ([xshift = 2cm]t) (nn){
      \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.3,shift={+(-1,2.5)}}]
       \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
        \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
        \node[rectangle] at (1,1) {};
        \foreach \name / \y in {1,...,4}
            \node[neuron] (I-\name) at (0,-\y) {};
        \foreach \name / \y in {1,...,3}
            \node[neuron] (H-\name) at (1,-\y-0.5) {};
        \foreach \name / \y in {1,...,4}
            \node[neuron] (O-\name) at (2,-\y) {};
        \foreach \source in {1,...,4}
            \foreach \dest in {1,...,3}
                \draw [-latex] (I-\source) -- (H-\dest);
        \foreach \source in {1,...,3}
            \foreach \dest in {1,...,4}
                \draw [-latex] (H-\source) -- (O-\dest);
      \end{tikzpicture}
    };
    \node[align = center, text width = 1cm] at ([yshift = 0.7cm]nn.north) {\baselineskip=0pt \Small Recog. model\par};
    \draw [->] (t.east) -- ([xshift = -0.5cm]nn.west);

    \node[draw,rounded corners, inner sep = 10] at ([xshift = 4.2cm,yshift = -1cm]) (s){Search};
%    \node at ([xshift=-7pt,yshift=5pt]s.north west) {$\mathcal{D}$};

    \draw [->] ([xshift = 0.5cm]nn.east) -- ([yshift = -0.25cm]s.west);
    \draw [->,rounded corners] (d.east) -- ([yshift = 2cm]nn.center) -- ([yshift = 0.25cm]s.west);

    \node[align=left] at (7,-1) (f) {Programs};
    \draw [->  ] (s.east) -- (f.west);

    \draw [->  ,rounded corners] (t.south) -- ([yshift = -0.5cm]t.south) -- ([yshift = -0.5cm] s.south |- t.south) -- (s.south);
  \end{tikzpicture}}
  %% Recognition model: Neural net.\\
  %% Red: Bottom-up inference.\\
  %% Gray: Observed.\\
  %% White: Latent.\\
  %% Boxed (plate): Repeated.\\

\end{frame}
\begin{frame}[t]{Sleep-G: Memory consolidation}
  \includegraphics[width = 11cm]{figures/animation/Sleep-G.eps}
  \vfill
  \textbf{Fragment Grammars: O'Donnell 2015.}\\
  Orange: Code fragments.\\
  %% Gray: Observed.\\
  %% White: Latent.\\
  %% Boxed (plate): Repeated.\\

\end{frame}
\begin{frame}[t]{Sleep-R: Objective}
  \includegraphics[width = 11cm]{figures/animation/Wake.eps}

    \vfill
    Recognition model predicts distribution over program, conditional on task.\\
    Training: (program, task) pairs\\
    Objective: Predict program w/ (1) high prior under DSL \& \\(2) high likelihood for task

\end{frame}
\begin{frame}[t]{Sleep-R: Experience replay}
  \includegraphics[width = 11cm]{figures/animation/Sleep-R-1.eps}

    \textbf{Train on (program, task) pairs found during waking}
\end{frame}
\begin{frame}[t]{Sleep-R: Dreaming}
  \includegraphics[width = 11cm]{figures/animation/Sleep-R-2.eps}

    \textbf{Train on (program, task) pairs sampled from DSL}

@@

@@

  Discovers 38 concepts, including `filter'
\end{frame}

\begin{frame}{List functions: Learning curves on hold out tasks}

  \begin{center}
      \includegraphics[width = 6.5cm]{figures/listLearningCurve.eps} 
    \end{center}

@@

  \begin{block}{Tasks}
    \texttt{task : unit -> image}
  \end{block}

@@

  \textbf{Takeaway:}
  \begin{itemize}
    \item Humans flexibly adapt to diverse sets of new problem domains --- DreamCoder takes a step in this direction
  \end{itemize}
  \pause

  %% \textbf{Future work}

  %% \begin{itemize}
  %%   \pause
  %% \item Theory learning
  %%   \item Planning
  %%   \end{itemize}

    \pause{}

    \begin{figure}
        \includegraphics[scale=0.44]{figures/xkcd_star.png}
        \caption*{Thanks to \textsc{xkcd} we can rule out $36^\circ$ as a primitive.}
    \end{figure}

>>>>

@@

\end{frame}
\begin{frame}[t]{DSL learning as Bayesian inference}
  %  \includegraphics[width = 11cm]{figures/animation/EC.eps}
\centering  \begin{tikzpicture}[scale=1.3,line width=0.5mm]

@@

\pause

\textbf{New:} amortized inference +\\ better program representation (Lisp) + \\better DSL inference

\end{frame}
\begin{frame}{Wake --- as in Exploration/Compression Algorithm}
  \only<1>{
    \includegraphics[width=11cm]{figures/teachDC/.inkslides-ec/slide-0.pdf}
  }
  \only<2>{
    \includegraphics[width=11cm]{figures/teachDC/.inkslides-ec/slide-1.pdf}
  }
  \only<3>{
    \includegraphics[width=11cm]{figures/teachDC/.inkslides-ec/slide-2.pdf}
  }
  \only<4>{
    \includegraphics[width=11cm]{figures/teachDC/.inkslides-ec/slide-3.pdf}
  }
  \only<5>{
    \includegraphics[width=11cm]{figures/teachDC/.inkslides-ec/slide-4.pdf}
  }

@@

\begin{frame}{DreamCoder --- Sleep-G}
  \only<1>{
    \includegraphics[width=11cm]{figures/teachDC/.inkslides-sleep-g/slide-4.pdf}
  }
\end{frame}
\begin{frame}{DreamCoder --- Sleep-R (\textbf{Experience Replay})}
  \only<1>{
    \includegraphics[width=11cm]{figures/teachDC/.inkslides-sleep-r-1/slide-4.pdf}
  }
\end{frame}
\begin{frame}{DreamCoder --- Sleep-R (\textbf{Dreaming})}
  \only<1>{
    \includegraphics[width=11cm]{figures/teachDC/.inkslides-sleep-r-2/slide-4.pdf}
  }
\end{frame}

%% \begin{frame}[t]{Wake: Problem solving}
%%   \includegraphics[width = 11cm]{figures/animation/Wake.eps}

%%   \vfill
%%   \only<1>{
%%   Recognition model: Neural net.\\
%%   Red: Bottom-up inference.\\
%%   Gray: Observed.\\
%%   White: Latent.\\
%%   Boxed (plate): Repeated.\\
%%   }
%%   \only<2>{
%%   \begin{tikzpicture}[xscale=1.25,line width=1mm]
%%     \node at (0,0) (d){DSL};
%%     \node at ([yshift = -2cm]d) (t){$\text{Task}$};

%%     \node at ([xshift = 2cm]t) (nn){
%%       \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.3,shift={+(-1,2.5)}}]
%%        \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
%%         \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
%%         \node[rectangle] at (1,1) {};
%%         \foreach \name / \y in {1,...,4}
%%             \node[neuron] (I-\name) at (0,-\y) {};
%%         \foreach \name / \y in {1,...,3}
%%             \node[neuron] (H-\name) at (1,-\y-0.5) {};
%%         \foreach \name / \y in {1,...,4}
%%             \node[neuron] (O-\name) at (2,-\y) {};
%%         \foreach \source in {1,...,4}
%%             \foreach \dest in {1,...,3}
%%                 \draw [-latex] (I-\source) -- (H-\dest);
%%         \foreach \source in {1,...,3}
%%             \foreach \dest in {1,...,4}
%%                 \draw [-latex] (H-\source) -- (O-\dest);
%%       \end{tikzpicture}
%%     };
%%     \node[align = center, text width = 1cm] at ([yshift = 0.7cm]nn.north) {\baselineskip=0pt \Small Recog. model\par};
%%     \draw [->] (t.east) -- ([xshift = -0.5cm]nn.west);

%%     \node[draw,rounded corners, inner sep = 10] at ([xshift = 4.2cm,yshift = -1cm]) (s){Search};
%% %    \node at ([xshift=-7pt,yshift=5pt]s.north west) {$\mathcal{D}$};

%%     \draw [->] ([xshift = 0.5cm]nn.east) -- ([yshift = -0.25cm]s.west);
%%     \draw [->,rounded corners] (d.east) -- ([yshift = 2cm]nn.center) -- ([yshift = 0.25cm]s.west);

%%     \node[align=left] at (7,-1) (f) {Programs};
%%     \draw [->  ] (s.east) -- (f.west);

%%     \draw [->  ,rounded corners] (t.south) -- ([yshift = -0.5cm]t.south) -- ([yshift = -0.5cm] s.south |- t.south) -- (s.south);
%%   \end{tikzpicture}}
%%   %% Recognition model: Neural net.\\
%%   %% Red: Bottom-up inference.\\
%%   %% Gray: Observed.\\
%%   %% White: Latent.\\
%%   %% Boxed (plate): Repeated.\\

%% \end{frame}
%% \begin{frame}[t]{Sleep-G: Memory consolidation}
%%   \includegraphics[width = 11cm]{figures/animation/Sleep-G.eps}
%%   \vfill
%%   \textbf{Fragment Grammars: O'Donnell 2015.}\\
%%   Orange: Code fragments.\\
%%   %% Gray: Observed.\\
%%   %% White: Latent.\\
%%   %% Boxed (plate): Repeated.\\

%% \end{frame}
%% \begin{frame}[t]{Sleep-R: Objective}
%%   \includegraphics[width = 11cm]{figures/animation/Wake.eps}

%%     \vfill
%%     Recognition model predicts distribution over program, conditional on task.\\
%%     Training: (program, task) pairs\\
%%     Objective: Predict program w/ (1) high prior under DSL \& \\(2) high likelihood for task

%% \end{frame}
%% \begin{frame}[t]{Sleep-R: Experience replay}
%%   \includegraphics[width = 11cm]{figures/animation/Sleep-R-1.eps}

%%     \textbf{Train on (program, task) pairs found during waking}
%% \end{frame}
%% \begin{frame}[t]{Sleep-R: Dreaming}
%%   \includegraphics[width = 11cm]{figures/animation/Sleep-R-2.eps}

%%     \textbf{Train on (program, task) pairs sampled from DSL}

@@

  Discovers 38 concepts, including `filter'. With suitable curriculum can also learn `map', `fold', etc.
\end{frame}

\begin{frame}{List functions \& Text editing: Learning curves on hold out tasks}

  \begin{center}
    \includegraphics[width = 5cm]{figures/listLearningCurve.eps}
\hfill    \includegraphics[width = 5cm]{figures/textLearningCurve.eps} 
    \end{center}

@@

  \begin{block}{Tasks}
    \texttt{task : image}
  \end{block}

@@

   \underline{More human-like machine intelligence:}\\Flexibly adapting to new problem domains by
   acquiring a domain-specific language-of-thought (DSL), while jointly learning how to use the language of thought (recognition model)

  \pause

\hspace{-1cm}\includegraphics[width = 12cm]{figures/finale.png} 
  \end{frame}
